The population of long-period exoplanets I. --- \\ Fully-automated discovery \& characterization of long-period transiting exoplanet candidates in the \kepler\ archive

### 2016-02-27 document/ms.tex
finally writing some words

\documentclass[twocolumn]{aastex6}
% \documentclass[onecolumn]{aastex6}
% \documentclass[preprint]{aastex}

@@

\newcommand{\T}{\ensuremath{\mathrm{T}}}
\newcommand{\dd}{\ensuremath{\,\mathrm{d}}}
\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}
\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
\newcommand{\densityunit}{{\ensuremath{\mathrm{nat}^{-2}}}}

@@

% Notation for this paper.
\newcommand{\period}{{\ensuremath{P}}}
\newcommand{\unit}[1]{{\ensuremath{\,\mathrm{#1}}}}

\newcommand{\params}{{\ensuremath{\bvec{w}}}}
\newcommand{\poppars}{{\ensuremath{\theta}}}

@@

\title{%
Single \& double transits discovered in the \kepler\ archive: \\
the frequency of long-period exoplanets
}

@@

There are two main technical barriers to a search for single transit events.
The first is that the transit probability for long-period planets is very low;
scaling as $\propto\period^{-5/3}$ for orbital periods longer than the
baseline of contiguous observations.
Therefore, even if long-period planets are intrinsically common, they will
still be underrepresented in a transiting sample.
The second challenge is that there are substantial signals in the observed
light curves caused by stochastic processes---both instrumental
(pointing jitter, temperature variations, \etc) and astrophysical (stellar
variability, \etc)---that can masquerade as transit signals.
In practice, even using the most sophisticated systematics removal methods,
these false signal far outnumber the true single transits.

@@

\section{The fully-automated search for long-period transiting exoplanets}

@@

% The search for long-period planets presents several unique challenges.
% Of these, the most important is the inevitable sensitively to signals falsely
% identified as being caused by a transit exoplanet when it is actually a
% systematic false alarm (noise) or an astrophysical false positive.
% In the following paragraphs, we detail our method for addressing these
% challenges and quantifying their effects.
% In summary, we identify false alarms by developing a conservative automated
% candidate vetting algorithm using the full light curve of each target star and
% measure the incompleteness introduced by this procedure.
% To account for astrophysical false positives, we discard candidates that fail
% the standard transit centroid tests and then statistically incorporate
% uncertainties introduced by false positives missed by this procedure at a
% population level.

@@

>>>>

% \documentclass[twocolumn]{aastex6}
\documentclass[onecolumn]{aastex6}
% \documentclass[preprint]{aastex}

@@

\newcommand{\T}{\ensuremath{\mathrm{T}}}
\newcommand{\dd}{\ensuremath{\,\mathrm{d}}}
\newcommand{\unit}[1]{{\ensuremath{\,\mathrm{#1}}}}
\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}
\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
\newcommand{\densityunit}{{\ensuremath{\mathrm{nat}^{-2}}}}

@@

% Notation for this paper.
\newcommand{\meanpars}{{\ensuremath{\bvec{\theta}}}}
\newcommand{\kernpars}{{\ensuremath{\bvec{\alpha}}}}

\newcommand{\params}{{\ensuremath{\bvec{w}}}}
\newcommand{\poppars}{{\ensuremath{\bvec{\beta}}}}

@@

\title{%
The population of long-period exoplanets I. --- \\
Fully-automated discovery \& characterization of
long-period transiting exoplanet candidates in the \kepler\ archive
}

@@

There are two main technical barriers to a search for single transit events.
The first is that the transit probability for long-period planets is very low;
scaling as $\propto P^{-5/3}$ for orbital periods longer than the
baseline of contiguous observations.
Therefore, even if long-period planets are intrinsically common, they will
still be underrepresented in a transiting sample.
The second challenge is that there are substantial signals in the observed
light curves caused by stochastic processes---both instrumental
(pointing jitter, temperature variations, \etc) and astrophysical (stellar
variability, \etc)---that can masquerade as transit signals.
In practice, even using the most sophisticated systematics removal methods,
these false signal far outnumber the true single transits.

@@

\section{A fully-automated search method}

@@

@@

In this step of the method, the goal is to discard any signals that are not
sufficiently ``transit-like'' in shape.
To quantify this decision, we perform a model comparison between a physical
transit model and a set of other parameterized models for systematics.
In order for a candidate to pass this vetting step, the transit model must be
``preferred'' to any other model as measured using the Bayesian Information
Criterion (BIC).
The BIC is not the optimal choice for this model comparison but it is
computationally intractable to compute thousands of precise marginalized
likelihoods for each model.
The BIC can be efficiently computed and it exhibits the desired
behavior---increasing with the likelihood but flexible models are
penalized---and we find that it performs sufficiently well in practice.

For up to three candidate transit times per light curve, we select a
contiguous chunk of light curve approximately centered on the proposed transit
with no more than 500 cadences \dfmtodo{check me} and compute the BIC of each
model for this data set.
We will define the BIC for a model $k$ in the set of $K$ models is given by
\begin{eqnarray}
\mathrm{BIC}_k &=& \ln \mathcal{L}^* - \frac{J}{2}\,\log N
\end{eqnarray}
where the likelihood function $\mathcal{L}$ is evaluated at its maximum, $J$
is the number of free parameters in the model, and $N$ is the number of
data points in the data set.

For each model, we describe the data using a Gaussian Process (GP;
\dfmtodo{cite}) with a Mat\'ern-3/2 covariance and mean given by the chosen
model $m_k(t;\,\meanpars)$ parameterized by the parameter vector \meanpars.
This generative model is specified by the following likelihood function:
\begin{eqnarray}
\mathcal{L} = \ln p_k(\bvec{y}\,|\,\meanpars,\,\kernpars) &=&
- \frac{1}{2}\,\bvec{r}_k(\meanpars)^\T\,K(\kernpars)^{-1}\,
    \bvec{r}_k(\meanpars)
- \frac{1}{2}\log\det K(\kernpars) - \frac{N}{2} \log{2\,\pi}
\end{eqnarray}
where \bvec{y} is the vector of flux measurements, \bvec{t} are the
corresponding times, and
\begin{eqnarray}
\bvec{r}_k(\meanpars) &=& \bvec{y} - m_k(\bvec{t};\,\meanpars) \quad.
\end{eqnarray}
The covariance matrix $K(\kernpars)$ is parameterized by the parameters
$\kernpars = (\alpha,\,\tau)$ and its elements are given by
\begin{eqnarray}
\left[ K(\kernpars) \right]_{ij} &=& \sigma_i^2\,\delta_{ij}
    + \alpha^2 \left[ 1+\frac{|t_i - t_j|}{\sqrt{3}\,\tau} \right]
      \exp \left(-\frac{|t_i - t_j|}{\sqrt{3}\,\tau}\right)
\end{eqnarray}
where $\sigma_i$ is the reported uncertainty on the $i$-th flux measurement.

The gradients of the likelihood with respect to the parameters \meanpars\
and \kernpars\ are given by
\begin{eqnarray}
\frac{\dd\ln p_k(\bvec{y}\,|\,\meanpars,\,\kernpars)}{\dd \meanpars} &=&
\frac{\dd m_k(\bvec{t};\,\meanpars)}{\dd\meanpars}^\T \, K(\kernpars)^{-1} \,
    m_k(\bvec{t};\,\meanpars)
\end{eqnarray}
and
\begin{eqnarray}
\frac{\dd\ln p_k(\bvec{y}\,|\,\meanpars,\,\kernpars)}{\dd \kernpars} &=&
\frac{1}{2}\,\mathrm{Tr}\left(
    \left[ \bvec{\phi}\,\bvec{\phi}^\T - K(\kernpars)^{-1} \right]
    \,\frac{\dd K(\kernpars)}{\dd\kernpars}
\right)
\end{eqnarray}
where
\begin{eqnarray}
\bvec{\phi} &=& K(\kernpars)^{-1}\,\bvec{r}_k(\meanpars) \quad.
\end{eqnarray}
Therefore, if we can compute the gradient of the model with respect to its
parameters, we can efficiently optimize this function to find the maximum
likelihood parameters using a gradient-based non-linear optimization method.
To compute the gradients of the transit model, we use a compile-time automatic
differentiation library \dfmtodo{cite Ceres} and for the other models, we
choose models that can be easily differentiated analytically.

The models are:
\begin{itemize}
{\item a limb-darkened, exposure time integrated transit light curve,}
{\item a pure Gaussian Process model to capture stellar variability,}
{\item a single outlier,}
{\item an step function, and}
{\item a box.}
\end{itemize}
